% Reduced-Order Models

Often an engineering team seeks to perform optimization and calibration routines on a set of computer codes which may require several hours, if not days, to complete a single simulation. Given the limited computational budget available to such teams a surrogate model is typically sought for the computational codes. The use of a surrogate model for an expensive computer code effectively exchanges predictive accuracy for simulation execution time. Unfortunately, to construct a surrogate model the expensive computer code must still be executed a certain number of times, although still significantly less times than required to conduct an optimization study. At this point there are two directions that can be taken to construct the surrogate model. If a computational budget is limited to say, $N$ evaluations of the objective function then it is necessary to somehow optimize the set of input points at which the objective computer code is evaluated. The general strategy for constructing a surrogate is outlined in Fig. \ref{fig:surrogate_process}, with each component described in detail in the proceeding sections. Sampling plan optimization strategies are described in section \ref{sec:sampling_plans}, while surrogates based on the resulting sampling plan, namely Kriging, is described in section \ref{sec:kriging}. If there is more flexibility in the computational budget then collocation-based surrogates should be considered since they offer the benefits of built-in convergence metrics.        

Computer codes that model physical phenomena typically accept an input file whose purpose is to describe specific conditions in the universe being modeled by the code. The code is executed and the affects of the conditions on some dependent quantities are output. From this perspective computer codes can be viewed and treated in much the same way functions mathematicians deal with are treated. Like matrices, functions with certain properties can undergo various decompositions that offer insight into their structure. The purpose of section \ref{sec:func_decomp} is to describe a technique for decomposing functions into orthogonal components, with the ultimate intention of applying the technique to computer codes. It is hoped that the decomposition of the computer code in terms of its inputs reveals which inputs play the most active roles in the underlying physics. Keeping only the most active dimensions in the decomposition, a reduced order model is effectively built. However, the function decomposition technique described in section \ref{sec:func_decomp} describes only half the story. To create a reduced order model of a computer code that can be efficiently evaluated at any state point in the original parameter space an efficient, multidimensional, interpolation scheme is needed. Such a scheme is discussed in section \ref{sec:smolyak_sg}. The coupling between interpolation and function decomposition that creates a usable reduced order model is then described in section \ref{sec:decomp_and_interp}.    
\begin{figure}
\caption{\label{fig:surrogate_process}
General flow diagram for constructing a surrogate for an expensive computer code.}
\begin{tikzpicture}[node distance = 2.5cm, auto]
\tikzstyle{every node}=[font=\small]
\node[block, label=right:(Define computer experiments)] (samp_plan) {Sampling Plan};
\node[block, below of= samp_plan, label=right:(Quantitative evaluation of designs)] (obs) {Observations};
\node[block, below of= obs, label=right:(Strategic processing of design variables)](dim_red) {Dimension Reduction};
\node[block, below of= dim_red, label=right:(Create fast mapping)](surr) {Surrogate Construction};
\node[block, below of= surr, label=right:(Evaluate surrogate many times)](opt) {Optimization};
\path [line] (samp_plan.south) -- (obs.north);
\path [line] (obs.south) -- (dim_red.north);
\path [line] (dim_red.south) -- (surr.north);
\path [line] (surr.south) -- (opt.north);
\end{tikzpicture}  
\end{figure}

% Optimized Sampling Plans
\input{./Chapter2/optimized_sampling_plan}
% Kriging
\input{./Chapter2/kriging}
% Function Decompositions
\input{./Chapter2/function_decompositions}
% Smolyak Sparse Grids
\input{./Chapter2/smolyak_grids}
% Combining Decomposition and Interpolation
\input{./Chapter2/decomposition_and_interpolation}
