In this section Bison is used to model the Ris\o~ AN3 power ramp fission gas release time series. More specifically, given the wide uncertainty in the input fission gas release parameters the problem of identifying parameter values resulting in Bison predictions most similar to experimental data is tackled. Since a single Bison simulation of the Ris\o~ AN3 power ramp is a computationally expensive, the surrogate construction techniques described in chapter \ref{chap:rom} will be employed. Recall that calibration exercises require thousands of instances of computer code simulations. Consequently, without the use of surrogates parameter calibration would not be feasible in this problem. Of the two surrogate construction techniques described, namely Kriging and the collocation approach, Kriging is more applicable for the problem in hand. Both surrogate techniques are designed for modeling surrogates of scalar quantities. However, in this problem a surrogate for an entire time series is desired. Of course, a surrogate can be constructed at each time-step but this would be extremely expensive considering there are $\mathcal{O}(100)$ time-steps. Kriging is used here mainly because it can most easily be extended when only a set and limited number of Bison simulations can be afforded. 

\subsection{Uncertain Parameters}
\label{subsec:uncertain_params}
\input{./Chapter4/uncertain_params}

\subsection{Principal Component Analysis}
\label{subsec:fgrPCA}
\input{./Chapter4/pca}

\subsection{Time-Series Surrogate}
\label{subsec:ts_surrogate}
\input{./Chapter4/time_series_surrogate}


%Want to find which set of fgr parameters will cause bison to produce a time series closely as possible to the experiment as possible. To this end, need viable range of parameter values and parameters to tweak. Giovaani and Laura have done extensive research to obtain such quantities. The fgr parameters that will be tweaked are... and their uncertaineis are... Show where they appear in FGR equations. 
%Kriging based on producing surrogates for scalar quantities. We want a surrogate for time series. Effectively need to train surrogate at each time step. PCA to the rescue. We want to input some FGR parameters and get a time series out of it. Make sure same times are used, 150 points are taken for bison fgr and experimental data. Splines used. Tests for more and less were conducted... 150 points adeuqate to not differ from original data.  
%Doesn't make sense to use collocation here. 
%Only 3 kriging surrogates
%error metric is MSE. others were tried.
%cross-validation of surrogates
%smoothing results?
