\section{Uncertainties in Nuclear Data}
\label{sec:cross_sec_uncertainties}

\subsection{Cross Section Uncertainty Overview}
\label{subsec:xsec_uq_overview}

While the methods described in this thesis are applicable to a plethora of engineering applications, the target application here concerns the simulation of nuclear reactors. A main component of any reactor analysis is a description of the neutronics, the physical processes behind neutron transport inside a reactor. When studying uncertainty quantification and sensitivity analysis of the neutronics in some reactor system, the primary  player has been shown to be uncertainty in cross section data \cite{Khalik_Turinsky} \cite{Jessee_Khalik_Turinsky}. Cross section uncertainties for various reactions, energies, and nuclides are accumulated during their experimental determination. Many of the cross sections exhibit correlations among themselves that must be taken into account. A problem arises when one becomes interested in performing uncertainty and sensitivity analysis on a core simulator since the simulators use processed cross section data. The experimental cross section uncertainties and covariances must be carefully propagated down through the standard cross section processing regime.   

Until recently, the primary approach for cross section uncertainty propagation involved the application of first-order perturbation theory. The perturbation theory approach is described in detail in \cite{MLWilliams}. Basically, for neutronics uncertainty analysis this approach entails the solution of an adjoint transport equation for each response of interest, allowing for the efficient retrieval of sensitivity coefficients for all input data. The sensitivity coefficients can then be used to obtain response uncertainties by being operated on by the inputs' covariance matrix. However, the linear approximations in perturbation theory and generalized perturbation theory may fail in certain scenarios. In particular, if the ratio of neutron loss due to absorption is high, such as when a control rod is rapidly inserted, second order effects may become substantial \cite{MLWilliams}. In addition, while perturbation theory may be computationally efficient for problems where only a handful of responses are of interest, problems with many responses may be burdensome due to the need to solve the generalized adjoint transport equation for each response. Not to mention, implementation of the perturbation theory approach is intrusive to engineering codes which makes uncertainty quantification of coupled, multiphysics code systems  infeasible.         

With the recent increase in accessibility to parallel computing environments stochastic sampling methods have become popular since they do not apply any kind of approximations to the physics at hand. Consequently, for cross section uncertainty propagation sampling methods have also been adopted to replace perturbation theory. A description of the process for propagating experimental uncertainties in cross section data to few group cross sections is described in this section. The statistical sampling framework for cross section data is extremely flexible since uncertainties can be calculated for any few group parameters. The same cannot be said of perturbation theory, which must formulate responses as ratios of reaction rates. Consequently, uncertainties for few group transport cross sections and assembly discontinuity factors are difficult to obtain using generalized perturbation theory.   

\subsection{Sampling Method}
\label{subsec:xsec_sampling}

The method for producing stochastically sampled few-group cross sections will be described. Although the method described is quite general, to provide clarity it will be described in the context of the \ac{SCALE} code \cite{SCALE}. Specifically, \ac{SCALE}'s \ac{ENDF} libraries and cross section processing utilities will be detailed. In the \ac{ENDF} libraries the multigroup cross sections are assumed to be Gaussian and consequently, they can be fully characterized by their means and covariances. The \ac{SCALE} 44-group \ac{ENDF}/B-VII covariance matrix contains generic multigroup covariance data. For a problem where $m$ nuclide-reaction combinations are of interest the pertinent covariance matrix is expanded to a size of $\left[m\cdot 44\right]\times \left[m\cdot 44\right]$. The Gaussian assumption imposed on the cross section values implies the cross sections can take on negative values. Since cross sections physically cannot take on negative values their distributions are truncated \cite{Klein_Gallner}.      

The first step towards obtaining perturbed few-group cross sections is to sample the generic multigroup covariance library. The covariance data schema in \ac{SCALE} is given as relative values of infinitely dilute cross sections \cite{Williams_Ilas}. Denote the 1D energy dependent, pointwise cross section for reaction $x$ as $\sigma_x(E)$. The infinitely dilute cross section in energy group $g$ can then be calculated as,
\begin{equation}
\label{eq:infinitely_dilute_xsec}
   \sigma_{x,g}(\infty) = \frac{\langle\sigma_x(E)\rangle}
    {\Delta u_g}   
\end{equation}  
where the bracket notation indicates a lethargy inner product over the energy interval covered by group $g$. Consequently, when the multigroup covariance matrix in \ac{SCALE} is sampled the resulting perturbation $\sigma_{x,g}'$ to the infinitely dilute cross sections can be expressed in terms of a perturbation factor $Q_{x,g}$,
\begin{eqnarray}
\label{eq:infinitely_dilute_xsec_prime}
   \sigma_{x,g}'(\infty) &=& \left(1 + \frac{\Delta\sigma_{x,g}(\infty)}
    {\sigma_{x,g}(\infty)}\right)\sigma_{x,g}(\infty) \nonumber \\
     \nonumber \\
      &=& Q_{x,g}\sigma_{x,g}(\infty). 
\end{eqnarray}  
The perturbations in Eq. \ref{eq:infinitely_dilute_xsec_prime} will provide generic multigroup cross section values to propagate through a transport solver, which can then yield perturbed few-group cross section values. However, to make the perturbed cross sections problem specific they must undergo adjustments due to resonance self shielding. For this purpose, perturbed pointwise cross sections and perturbed Bondarenko self-shielding factors are required. The pointwise, or continuous energy, cross sections are needed to perform 1D transport calculations in the resolved resonance range while the Bondarenko factors are applicable in the unresolved energy range. Since the pointwise cross sections, Bondarenko factors, and generic infinitely dilute multigroup cross section are related algebraically it is necessary to perturb each in terms of the factor $Q_{x,g}$. 

In \cite{Williams_Ilas} the authors show that, given a perturbation factor $Q_{x,g}$, the appropriate pointwise cross sections perturbation can be obtained by,
\begin{equation}
\label{eq:pointwise_xsec_prime}
   \sigma_x'(E) = Q_{x,g}\sigma_x(E) \hspace{1cm} E\in g.
\end{equation} 
From Eq. \ref{eq:pointwise_xsec_prime} it is clear that the pointwise cross section data gets perturbed uniformly in a given energy group $g$, which is a valid approximation if the energy width is small. The authors in \cite{Williams_Ilas} note that although this approximation leads to discontinuities at the energy group boundaries, it does not impact the resulting multigroup data significantly. No significant impact is expected because the multigroup data is averaged using fluxes and cross sections from the same energy bin. 

At this point, infinitely dilute multigroup cross sections and pointwise cross sections can be consistently perturbed using the \ac{SCALE} \ac{ENDF} covariance library. To proceed with a transport calculation using perturbed parameters, perturbed Bondarenko self-shielding factors $f_{x,g}'(\sigma_0)$ are also needed. The Bondarenko factors are expressed in terms of the background cross section $\sigma_0$, pointwise cross sections, and infinitely dilute cross sections as,     
\begin{equation}
\label{eq:bondarenko}
   f_{x,g}(\sigma_0) = \frac{1}{\sigma_{x,g}(\infty)}
    \Big\langle \frac{\sigma_x(E)}{\sigma_t(E) + \sigma_0}\Big\rangle
     \Big/
     \Big\langle \frac{1}{\sigma_t(E) + \sigma_0}\Big\rangle.
\end{equation}
Substituting Eqs. \ref{eq:infinitely_dilute_xsec_prime} and \ref{eq:pointwise_xsec_prime} into Eq. \ref{eq:bondarenko}, the perturbed Bondarenko factor is obtained,
\begin{equation}
\label{eq:bondarenko_prime}
   f_{x,g}'(\sigma_0) = \frac{1}{Q_{x,g}\sigma_{x,g}(\infty)}
    \Big\langle \frac{Q_{x,g}\sigma_x(E)}{Q_{t,g}\sigma_t(E)+\sigma_0}\Big\rangle
     \Big/
     \Big\langle \frac{1}{Q_{t,g}\sigma_t(E) + \sigma_0}\Big\rangle.
\end{equation}
The authors in \cite{Williams_Ilas} show that Eq. \ref{eq:bondarenko_prime} can be evaluated by simply evaluating the unperturbed Bondarenko equation in Eq. 
\ref{eq:bondarenko} at a perturbed background cross section $\sigma_0' = \sigma_0/ Q_{t,g}$. Note that only those cross sections whose uncertainties are specified in the \ac{SCALE} covariance library can be processed and propagated through transport calculations. Since covariance data for 2D scattering distributions is not available in the \ac{SCALE} \ac{ENDF}/B-VII covariance library, their uncertainty affects cannot be quantified in any analyses \cite{Williams_Ilas}. 
 
\subsection{Cross Section Sampling in SCALE}
\label{subsec:scale_sampler}  

The method described in section \ref{subsec:xsec_sampling} for producing perturbed cross sections will be further described in the context of the modules in \ac{SCALE}. The first step is to produce perturbation factors using the module XSUSA by sampling the \ac{SCALE} \ac{ENDF}/B-VII covariance library. If a total of $N_1$ output files of transport solver are desired then $N_1$ sets of perturbation factors are to be obtained. Using the perturbation factors, the \ac{SCALE} module CLAROL+ yields perturbed, infinitely dilute, multigroup cross sections by essentially applying Eq. \ref{eq:infinitely_dilute_xsec_prime}. The resulting cross sections are then passed to the module CRAWDAD+, which outputs perturbed, continuous energy cross sections. To make the multigroup cross sections problem specific they are passed to the BONAMI module, which adjusts the cross sections to account for self-shielding effects in the unresolved energy region using the Bondarenko shielding factor method. Perturbed Bondarenko factors are used in BONAMI, as output from CLAROL+ upon application of Eq. \ref{eq:bondarenko_prime}.

At this point the cross sections still need to be processed for self-shielding effects in the resolved energy region. The \ac{SCALE} modules CENTRM/PMC are responsible for folding in these effects into the multigroup cross sections. CENTRM solves the neutron slowing down equation on an ultra fine spectra on the order of $\mathcal{O}(10^4)$ energy points. For the flux a linear interpolation scheme is used between energy points to produce a continuous energy solution over the entire energy range \cite{Williams_Gehin}. The CENTRM module incorporates actual composition, temperature, and environment dependent effects into the resonance shielded cross sections. PMC accepts the high resolution flux spectra output by CENTRM and averages pointwise nuclear data into multigroup cross sections, which are now tailored to solve the specific problem of interest.              

Finally, the problem specific multigroup cross sections output by CENTRM/PMC are sent to a lattice transport solver, which in this case is the \ac{SCALE} module NEWT. In general, NEWT can be replaced by any transport solver. However, for the purposes of producing few group cross sections a lattice transport solver will generate the multigroup flux distribution for some heterogeneous configuration, which is then used to generate homogenized few group cross sections and intra-assembly flux shapes. A script to extract desirable output quantities from the transport output, such as few group cross sections and kinetics parameters, is needed. Such a script is executed after each of $N_1$ calls to NEWT. Each call extracts a set of perturbed, few group cross sections that are ultimately propagated through a core simulator. The SAMPLER sequence in \ac{SCALE} automates the process of producing perturbed cross section sets \cite{Williams_Ilas}.      

With the availability of $N_1$ perturbed few group cross section sets, each set containing all the nuclear data required by a core simulator, the sets can be directly propagated through the core simulator. The desired results from each simulation can then be gathered and statistically analyzed in what is referred to as the 'one-step' method. However, in what is referred to as the 'two-step' method, the $N_1$ perturbed few group cross section sets can be used to create a few group covariance matrix \cite{TwoStep_Approach}. The few group covariance matrix is then sampled and each sample is propagated through the core simulator. Both the 'one-step' and 'two-step' methods have been shown to produce consistent results \cite{TwoStep_Approach}. However, the 'two-step' method has the advantage of breaking the limitation on the number of samples that can be propagated through the core simulator. In the 'one-step' method each core simulation requires a corresponding transport solution. Transport solutions are relatively expensive to compute when compared to core simulations. A flow diagram of the 'two-step' method is displayed in Figure \ref{fig:sampler_flow_diagram}. For the 'one-step' method $N_1=N_2$ whereas for the 'two-step' method $N_2 >> N_1$, allowing for the acquisition of better statistics.        
\begin{figure}
\caption['Two-Step' method flow diagram.]{\label{fig:sampler_flow_diagram}
Flow diagram of 'two-step' method for core simulators.}
\begin{tikzpicture}[node distance = 3.5cm, auto]
\tikzstyle{every node}=[font=\small]
   % Place nodes
   \node [block1] (xsusa) {\textbf{XSUSA}: Make random perturbations to covariance 
   library};
   \node [block, below of=xsusa] (clarol) {\textbf{CLAROL+}: 
   Perturb MG xsecs and shielding factors};
   \node [block, below of=clarol] (crawdad) {\textbf{CRAWDAD+}: 
   Perturb CE xsecs};
   \node [block, right of=clarol, node distance=8cm] (bonami) {\textbf{BONAMI}: 
   Perform self-shielding calculations in unresolved energy region};
   \node [block, right of=crawdad, node distance=8cm] (centrm) 
   {\textbf{CENTRM/PMC}: 
   Perform self-shielding calculations in resolved energy region};  
   \node [block, below of=centrm] (newt) {\textbf{NEWT}: 
   Perform transport calculations using perturbed xsecs};
   % Invisible nodes
   \node [left of= xsusa, node distance=2.5cm, scale=0.01] (xsusa_i) {};
   \node [left of= newt, node distance=4.5cm, scale=0.05] (newt_l) {};
   \node [block, below of=newt_l, node distance=3.0cm] (cov) {Sample 
   $C_{\Sigma_{FG}}$ matrix};
   \node [block, below of=cov] (core) {\textbf{Core Simulator}: Simulate perturbed
   input data};
   % Invisible npde
   \node [left of= cov, node distance=6.0cm, scale=0.05] (cov_i) {};
   \node [left of= core, node distance=3.5cm, scale=0.05] (core_i) {};
   \node [block1, below of=core_i, node distance=3.0cm] (stats) {Analyze $N_2$ 
   results};
   
   % Draw edges
   \path [line] (xsusa) -- node (xsusa_o) {$Q_{x,g}$} (clarol);
   \path [line] (clarol) -- (crawdad);   
   \path [line] (clarol) -- node (clarol_o) {$\sigma_{x,g}'(\infty)$, 
   $f_{x,g}'(\sigma_0)$} (bonami);
   \path [line] (crawdad) -- node (crawdad_o) {$\sigma_{x}'(E)$} (centrm);
   \path [line] (bonami) -- node[left] (bonami_o) {initial self-shielded data}
   (centrm);
   \path [line] (centrm) -- node[left] (centrm_o) {final self-shielded data} 
   (newt); 
   \path [-,draw] (newt) -| node[below]{Repeat $N_1$ times} (xsusa_i.north);
   \path [line]{} (xsusa_i.north) |- node{} (xsusa);  
   \path [line] (newt_l) -- node[left] (newt_lo){$C_{\Sigma_{FG}} =$ Cov 
   $(\Sigma_{FG_1}...\Sigma_{FG_{N_1}})$} (cov);
   \path [line] (cov) -- node[left] (cov_o){$\Sigma_{FG_i}'$} (core); 
   \path [-,draw] (core) -| node[below]{Repeat $N_2$ times} (cov_i.north);
   \path [line]{} (cov_i.north) |- node{} (cov);
   \path [line] (core_i) -- (stats); 
\end{tikzpicture}   
\end{figure}

